---
title: 'FMPH 291 Reproducibility: Assignment 2'
author: "Arnab Dey, Michael Cheung"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  slidy_presentation:
    highlight: tango
    slide_level: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

```{r load-packages, include=FALSE}
library(haven)
library(tidyverse)
library(summarytools)
library(compareGroups)
library(mediation)
library(broom)
library(broom.mixed)
library(lme4)
library(boot)
```

```{r load-dataset}

varlist_couples <- c("v001", "v020", "v021", "v024", "v312", "v025", "v012", "mv012", 
                     "v130", "v133", "mv133", "v205", "sm224", "s309",
                      "v202", "v203", "v204", "v205", "v191", "v602", "mv602",
                     "v614", "mv614")

df_master <- read_dta("C:/Users/arnab/OneDrive/Desktop/Analyses/Datasets/NFHS-4/Data/Couples Recode/IACR74DT/IACR74FL.dta", 
                      col_select = all_of(varlist_couples))
df_master <- as_factor(df_master)
```

```{r dataset for analysis, include=FALSE}
## All couples
df_couples <- df_master %>% 
  mutate(total_child = v202 + v203 + v204 + v205) %>% 
  filter(total_child != 0)
```

## Motivation for analysis

- Sex composition of children is known to reduce contraceptive use in India
- We want to understand if couple's desire for more children mediates the association 

## Dataset Used

- Demographic Health Survey - India conducted in 2015-16
- Couple's dataset where responses from both the male and the female partner are captured
- Total 63,696 cases of which 57,790 couples had at least 1 child
- We used the 57,790 cases for our analysis

## Key variables

* Outcome variable: Current use of any modern method of family planning by the couple (0/1)
* Exposure variable: Couples that have more girls than boys 
* Mediator: Couple's who are concordant in their desire for more children


## Methods

* We used Baron and Kenny's (1986) four step approach to understand the mediation effect
* Step-1 involved regressing the Exposure on the Outcome, adjusting for confounders
* Step-2 involved regressing the Exposure on the mediator
* Step-3 involved regressing the Mediator on the Outcome, adjusting for the exposure and other covariates
* We stratified the analysis by Total number of living children (less than 3 vs. 3 or more) 
* We also used the Mediation package in R to calculate the bootstrapped CIs (500 sims) around the ACME 


## Outcome of interest: Use of any modern contraceptive method by the couple

```{r create dependent variable for the analysis, echo = FALSE}

df_couples$any_modern <- with(df_couples, ifelse((v312 == "not using" | v312 == "rhythm/periodic abstinence" | 
                                               v312 == "withdrawal" | v312 == "lactational amenorrhea (lam)" | 
                                               v312 == "standard days"), 0, 1)) 

df_couples$any_modern <- with(df_couples, factor(any_modern, levels = c(0,1), labels = c("No or traditional methods", "Any modern method")))


df_couples %>% 
  group_by(any_modern) %>% 
  summarize(n = n()) %>% 
  mutate(pct = round(n/sum(n)*100, 2)) %>% 
  ggplot(aes(x = any_modern, y = pct)) +
  geom_col(fill = "#7a6374") + 
  geom_text(aes(label = pct), vjust = -0.5) +
  expand_limits(y = c(0,100)) +
  ylab("Percentage") +
  xlab("") + 
  theme_minimal()
```

## Exposure of interest: Couple's with more daughters than sons

```{r create the exposure variable, echo = FALSE}

df_couples <- df_couples %>% 
  mutate(more_daughters = ifelse(((v202 + v204) < (v203 + v205)), "More daughters than sons", "Less or equal no. of sons")) %>% 
  mutate(more_daughters = as.factor(more_daughters))

df_couples %>% 
  group_by(more_daughters) %>% 
  summarize(n = n()) %>% 
  mutate(pct = round(n/sum(n)*100, 2)) %>% 
  ggplot(aes(x = more_daughters, y = pct)) +
  geom_col(fill = "#7a6374") + 
  geom_text(aes(label = pct), vjust = -0.5) +
  expand_limits(y = c(0,100)) +
  ylab("Percentage") +
  xlab("") + 
  theme_minimal()

```

## Mediator variable: Couple's desire for more children

```{r create the mediator variable, echo = FALSE}
df_couples <- df_couples %>% 
  mutate(w_want_more = ifelse(as.numeric(v614) > total_child, 1, 0)) %>% 
  mutate(m_want_more = ifelse(as.numeric(mv614) > total_child, 1, 0)) %>% 
  mutate(mw_concord = ifelse(w_want_more == m_want_more, 1, 0)) %>% 
  mutate(both_want = ifelse((w_want_more == 1 & m_want_more == 1), 1, 0)) %>% 
  mutate(both_want_fac = factor(both_want, levels = c(0,1), labels = c("At least one partner does not want more children", "Both want more children")))

df_couples %>% 
  group_by(both_want_fac) %>% 
  summarize(n = n()) %>% 
  mutate(pct = round(n/sum(n)*100, 2)) %>% 
  ggplot(aes(x = both_want_fac, y = pct)) +
  geom_col(fill = "#7a6374") + 
  geom_text(aes(label = pct), vjust = -0.5) +
  expand_limits(y = c(0,100)) +
  ylab("Percentage") +
  xlab("") + 
  theme_minimal()

```

## Stratification by: Number of living children

```{r stratify by num child, echo = FALSE}
df_couples <- df_couples %>% 
  mutate(num_child = ifelse(as.numeric(total_child) > 4, 4, total_child)) %>% 
  mutate(num_child = factor(num_child, levels = c(1:4), labels = c("One child", "Two children", "Three children", "Four or more children")))
  
df_couples %>% 
  group_by(num_child) %>% 
  summarize(n = n()) %>% 
  mutate(pct = round(n/sum(n)*100, 2)) %>% 
  ggplot(aes(x = num_child, y = pct)) +
  geom_col(fill = "#7a6374") + 
  geom_text(aes(label = pct), vjust = -0.5) +
  expand_limits(y = c(0,100)) +
  ylab("Percentage") +
  xlab("") + 
  theme_minimal()

```

## List of confounding variables 

- Age of men and women (in years)
- Educational attainment of men and women (in years)
- Religion (Islam vs. others)
- Wealth Index (Range: 0 to 100)
- Place of residence (Rural vs. Urban)

```{r covariates}
## Education - Women
df_couples$w_edu_yrs <- as.numeric(df_couples$v133)

## Education - Men
df_couples$m_edu_yrs <- as.numeric(df_couples$mv133)

## Age of women
df_couples$w_age <- df_couples$v012

## Age of Men
df_couples$m_age <- df_couples$mv012

## Religion
df_couples$islam <- ifelse(df_couples$v130 == "muslim", 1,0)

## Wealth
# create a wealth index score based on this sample
df_couples$std_wealth_index = ((df_couples$v191 - min(df_couples$v191))/((max(df_couples$v191)-min(df_couples$v191)))*100)

## Place of residence
df_couples$urban <- df_couples$v025
```


```{r create datasets for stratified analysis}
## Couples with less than 3 children
df_couples_2 <- df_couples %>% 
  filter(total_child < 3)

## Couples with more than 2 children

df_couples_3 <- df_couples %>% 
  filter(total_child > 2)
```

# Mediation Analysis: couples with less than 3 children 

## Step-1: Effect of more daughters (Exposure) on Contraception Use (Outcome)
```{r Step-1,echo=TRUE}

model_1a <- glmer(any_modern ~ more_daughters + num_child + (1 | v001), data = df_couples_2, family = binomial("logit"))
tidy(model_1a, exponentiate = F, conf.int = T)[(1:2), c(3:4,8:9)]

```

## Step-2: Effect of more girls (Exposure) on Concordance between partners (mediator)

```{r Step-2, echo=TRUE}
model_1b <- glmer(both_want ~ more_daughters + (1 | v001), data = df_couples_2, family = binomial("logit"))
tidy(model_1b, exponentiate = F, conf.int = T)[(1:2), c(3:4,8:9)]
```

## Step-3: Effect of Concordance between partners (mediator) on Contraception Use (outcome)

```{r Step-3, echo=TRUE}
model_1c <- glmer(any_modern ~ both_want + more_daughters + w_edu_yrs + m_edu_yrs + w_age + m_age + 
                  islam + std_wealth_index + urban + num_child + (1 | v001), data = df_couples_2, family = binomial("logit"))
tidy(model_1c, exponentiate = F, conf.int = T)[(1:3), c(3:4,8:9)]
```

## Calculate NDE and NIE

```{r}
## Estimates
(model_1_NDE <- coef(summary(model_1c))[3,1])
(model_1_NIE <- coef(summary(model_1b))[2,1] * coef(summary(model_1c))[2,1]) 

## Confidence Intervals for NDE

#(model_1_NDE_ci <- confint(model_1c, parm = "more_daughtersMore daughters than sons", level = 0.95))

## Confidence Intervals for NDE
# Function to generate NIE

foo <- function(model_med, model_outcome, mediator, exposure) {
  model_nie <- coef(summary(model_1b))[2,1] * coef(summary(model_1c))[mediator,1]
  return(model_nie)
}

# bootstrap's bootstraps
set.seed(0112358)
library(boot)
boot_nie <- boot(df_couples_2, foo, R = 1000)
boot.ci(boot.out = boot_nie, type = c("basic"))


```


## Using the mediation package
```{r Mediation package, echo=TRUE}
model_1_med <- mediate(model_1b,model_1c, treat = "more_daughters", mediator = "both_want", boot = TRUE, sims = 500)
summary(model_1_med)
```

# Mediation Analysis: couples with 3 or more children

## Step-1: Effect of more daughters (Exposure) on Contraception Use (Outcome)
```{r More than 2 children: step-1, echo=TRUE}

model_2a <- glmer(any_modern ~ more_daughters + num_child + (1 | v001), data = df_couples_3, family = binomial("logit"))

tidy(model_2a, exponentiate = F, conf.int = T)[(1:2), c(3:4,8:9)]

```

## Step-2: Effect of more girls (Exposure) on Concordance between partners (mediator)

```{r, echo=TRUE}
model_2b <- glmer(both_want ~ more_daughters + (1 | v001), data = df_couples_3, family = binomial("logit"))
tidy(model_2b, exponentiate = F, conf.int = T)[(1:2), c(3:4,8:9)]
```

## Step-3: Effect of Concordance between partners (mediator) on Contraception Use (outcome)

```{r mediation for more than 2 children: step-3, echo=TRUE}
model_2c <- glmer(any_modern ~ both_want + more_daughters + w_edu_yrs + m_edu_yrs + w_age + m_age + 
                  islam + std_wealth_index + urban + num_child + (1 | v001) , data = df_couples_3, family = binomial("logit"))

tidy(model_2c, exponentiate = F, conf.int = T)[(1:3), c(3:4,8:9)]
```


## Using the mediation package
```{r, echo=TRUE}
model_2_med <- mediate(model_2b,model_2c, treat = "more_daughters", mediator = "both_want", boot = TRUE, sims = 500)
summary(model_2_med)
```

## Next steps / Questions

* Results from Mediation package do not match our results from product method  
* Undertake multi-level models
* How should we deal with interaction between the Exposure and the Mediator?
* When do we need sensitivity analysis?
* How can we incorporate multinomial outcomes in mediation analysis
